---
title: "A-21077-2021-09-23"
author: "by 21077"
output:
  html_document: default
---

## Question
Exercises 5.4, 5.9, 5.13, and 5.14 (pages 149-151, Statistical Computating with R).

## Answer

```{r include=FALSE}
set.seed(1234)
```

### 5.4

Write a function to compute a Monte Carlo estimate of the Beta(3, 3) cdf, and use the function to estimate F(x)for x =0.1, 0.2,... , 0.9. Compare the estimates with the values returned by the pbeta function in R.

```{r}
beta_3_3_pdf <- function(x){
  x^2*(1-x)^2/beta(3,3) 
}

beta_3_3_cdf <- function(x){
  m <- 10000
  u <- runif(m)
  mean(beta_3_3_pdf(u*x))*x
}
```

```{r}
est <- numeric(9)
true <- numeric(9)
for(i in c(1:9)){
  true[i] <- pbeta(i/10,3,3)
  est[i] <- beta_3_3_cdf(i/10)
}


knitr::kable(rbind(est,true,dif = est-true))  
```

### 5.9
Implement a function to generate samples from a Rayleigh(σ) distribution, using antithetic variables. What is the percent reduction in variance?

```{r}
rayleigh_pdf <- function(x,sigma = 1) {
  ifelse(x>0, x/sigma^2*exp(-x^2/2/sigma^2), 0)
}
rayleigh_cdf <- function(x, sigma = 1, R = 10000, antithetic = TRUE) {
  if(x<=0) return(0)
  
  u <- runif(R/2) 
  v <- ifelse(antithetic, 1-u, runif(R/2))
  u <- c(u, v)
  
  mean(rayleigh_pdf(x*u,sigma))*x
}
```

Verify the correctness of rayleigh_cdf with sigma = 1.

```{r}
x <- c(1:10) * 0.3
true <- numeric(10)
anti_MC <- numeric(10)
norm_MC <- numeric(10)

for(i in 1:10){
  true[i]    <-  integrate(rayleigh_pdf,0,x[i],abs.tol = 1e-5)$value
  anti_MC[i] <-  rayleigh_cdf(x[i],antithetic = TRUE)
  norm_MC[i] <-  rayleigh_cdf(x[i],antithetic = FALSE)
}

knitr::kable(rbind(x,true,anti_MC,norm_MC))
```

compute the percent reduction in variance of rayleigh_cdf(x = 1, sigma = 1).

```{r}
R <- 100
anti_MC <- numeric(R)
norm_MC <- numeric(R)

for(i in 1:R){
  anti_MC[i] <-  rayleigh_cdf(1,antithetic = TRUE)
  norm_MC[i] <-  rayleigh_cdf(1,antithetic = FALSE)
}

v1 <- var(anti_MC)
v2 <- var(norm_MC)
cat("Var of antithetic method is ",v1,'.\n')
cat("Var of normal method is ",v2,'.\n')
cat("The percent reduction in variance is ",(v2- v1)/v2*100,"%.\n")

```

Generate samples by using cdf.

```{r}
# Inverse transform algorithm 
rayleigh_sample <- function(n,sigma = 1, R = 10000, antithetic = TRUE){
  u <- runif(n)
  sam <- numeric(n)
  f <- function(x,c) rayleigh_cdf(x,sigma, R = R, antithetic = antithetic) - c
  
  for(i in c(1:n)){
    #sam[i] <- ifelse(f(1000,u[i])>0, uniroot(f,c(-0.1,1000),c=u[i])$root, 1000)
    sam[i] <- uniroot(f,c(-0.1,1000),c=u[i],extendInt = 'upX')$root
  }
  sam
}
```
```{r}
n <- 400
x <- rayleigh_sample(n)
x <- x[which(x<10)]
hist(x,prob=TRUE,ylim=range(0,1/exp(-0.5)), main = "Rayleigh with sigma = 1")
y <- seq(0,floor(max(x))+1,0.01)
lines(y,rayleigh_pdf(y))
```


### 5.13

Find two importance functions f1 and f2 that are supported on {x>1} and are ‘close’ to g(x).

```{r}
g <- function(x){
  x^2*exp(-x^2/2)/(2*pi)^0.5
}
```

I choose normal distribution with mean=0 and sd=1 restricted in $(1,\infty)$ as f1 and chi-square distribution with df=6 restricted in $(1,\infty)$ as f2.

```{r}
f1 <- function(x){
  dnorm(x)/(1-pnorm(1))
}
f1_sample <- function(n){
  sam <- numeric(n)
  for(i in 1:n){
    sam[i] <- rnorm(1)
    while(sam[i]<=1){
      sam[i] <- rnorm(1)
    }
  }
  return(sam)
}
f2 <- function(x){
  dchisq(x,6)/(1-pchisq(1,6))
}
f2_sample <- function(n){
  sam <- numeric(n)
  for(i in 1:n){
    sam[i] <- rchisq(1,6)
    while(sam[i]<=1){
      sam[i] <- rchisq(1,6)
    }
  }
  return(sam)
}
```

Which of your two importance functions should produce the smaller variance?

I think f1 importance function produce the smaller variance, because $\frac{g(x)}{f_1(x)} \sim x^2$ is polynomial and $\frac{g(x)}{f_2(x)}\sim e^{\frac{x(1-x)}{2}}$ is exponential.
This can be verified in the next question.

### 5.14

Obtain a Monte Carlo estimate by importance sampling.

```{r}
R <- 100
int_f1 <- numeric(R)
int_f2 <- numeric(R)

for(i in 1:R){
  n <- 100
  x1 <- f1_sample(n)
  x2 <- f2_sample(n)
  
  int_f1[i] <- mean(g(x1)/f1(x1))
  int_f2[i] <- mean(g(x2)/f2(x2))
}

cat("Use f1, the estimate is",mean(int_f1),"and the variance is",var(int_f1))
cat("Use f2, the estimate is",mean(int_f2),"and the variance is",var(int_f2))

```




