---
title: "A-21077-2021-09-23"
author: "by 21077"
output:
  html_document: default
---

## Question
Exercises 5.4, 5.9, 5.13, and 5.14 (pages 149-151, Statistical Computating with R).
## Answer
```{r include=FALSE}
set.seed(123)
```

### 5.4
Write a function to compute a Monte Carlo estimate of the Beta(3, 3) cdf, and use the function to estimate F(x)for x =0.1, 0.2,... , 0.9. Compare the
estimates with the values returned by the pbeta function in R.
```{r}
beta_3_3_pdf <- function(x){
  x^2*(1-x)^2/beta(3,3) 
}

beta_3_3_cdf <- function(x){
  m <- 10000
  u <- runif(m)
  mean(beta_3_3_pdf(u*x))*x
}
```

```{r}
est <- numeric(9)
true <- numeric(9)
for(i in c(1:9)){
  true[i] <- pbeta(i/10,3,3)
  est[i] <- beta_3_3_cdf(i/10)
}


knitr::kable(rbind(est,true,dif = est-true))  
```

### 5.9
Implement a function to generate samples from a Rayleigh(σ) distribution, using antithetic variables. What is the percent reduction in variance?

```{r}
rayleigh_pdf <- function(x,sigma) {
  ifelse(x>0, x/sigma^2*exp(-x^2/2/sigma^2), 0)
}
rayleigh_cdf <- function(x, sigma = 1, R = 1000, antithetic = TRUE) {
  if(x<=0) return(0)
  
  u <- runif(R/2) 
  v <- ifelse(antithetic, 1-u, runif(R/2))
  u <- c(u, v)
  
  mean(rayleigh_pdf(x*u,sigma))*x
}
```
```{r}
# Inverse transform algorithm 
rayleigh_sample <- function(n,sigma = 1, R = 1000, antithetic = TRUE){
  u <- runif(n)
  sam <- numeric(n)
  f <- function(x,c) rayleigh_cdf(x,sigma, R = R, antithetic = antithetic) - c
  
  for(i in c(1:n)){
    sam[i] <- ifelse(f(1000,u[i])>0, uniroot(f,c(-0.1,1000),c=u[i])$root, 1000)
  }
  sam
}
```




### 5.13

### 5.14